\vspace{24pt} Comparison of Different Data Layouts for Columnar In-Memory Databases on the Basis of an Example Application

### 2018-08-31 paper/paper.tex
intermediate commit

Since the data neither changes nor grows anymore, we decided to store the information in sorted order. As sorting key we utilized the trajectory identifier and the timestamp. In order to speed up window queries we materialized a unique identifier, which is the row ID. This ID simplifies the creation of a primary key. Otherwise, a primary key would have to include multiple columns, most likely all of them. The Hana database automatically creates an index on this column. Furthermore, we realized that timestamp operations perform relatively bad. In order to avoid timestamps we calculate the amount of seconds lasted and store those as an integer in the database. The speed improvement of this optimization will be examined in the next section (see section \ref{sec:bench_ts}).

@@

\subsubsection{timestamps}
\label{sec:bench_ts}
% Ich würde hierzu gerne noch nen Benchmark erstellen

@@

\subsubsection{Calculate distances}

>>>>

Since the data neither changes nor grows anymore, we decided to store the information in sorted order. As sorting key we utilized the trajectory identifier and the timestamp. In order to speed up window queries we materialized a unique identifier, which is the row ID. This ID simplifies the creation of a primary key. Otherwise, a primary key would have to include multiple columns, most likely all of them. The Hana database automatically creates an index on this column. Furthermore, we realized that timestamp operations perform relatively bad. In order to avoid timestamps we calculate the amount of seconds lasted and store those as an integer in the database.

@@

@@

\subsubsection{Calculate distances}
A key functionality of our software is the calculation of the profit per taxi. Therefore, we take three different factors into account:
\begin{enumerate}
    \item Taxi start: As soon as a passenger enters a taxi, he or she has to pay 11 Yuan
    \item Waiting time: There are several reasons why a vehicle has to wait (e.g. for the passenger). We assume, that this time is approxiamtely 10 percent of the ride. An hour of waiting time costs 48 Yuan.
    \item Distance driven: The price of a ride depends on the distance. A typical value is 2.40 Yuan in the city of Shenzhen.
\end{enumerate}
% Die Werte haben sich schon geändert ;)
The prices used for calculation are taken from a website.\footnote{\href{https://www.numbeo.com/taxi-fare/country\_result.jsp?country=China}{https://www.numbeo.com/taxi-fare/country\_result.jsp?country=China}} The values might have changed until now. However, prices will not change in a way, that our calculations become fundamentally wrong.

The calculation of aspect one and two is trivial since these are only simple additions or multiplication. Determining the distance is a more complex part, that we want to describe in detail now.

To calculate the distance we demand to do it completely in the database. When querying a full trajectory we have to use the sample point format, since any other layout depends on the python backend.

The simplest way to get the distance of a ride is the usage of spatial points (\textit{ST_POINTS}). This is a data type implemented in SAP Hana, that can be created out of a longitude and a latitude value. Furthermore, there is a function to compute the distance between those points. A single ride starts, when a passenger enters a taxi (occupancy changes from zero to one) and ends, when the passenger leaves (occupancy changes from one to zero). We sum up the lengths between all points sum uo all values for a whole trajectory.

We performed a first run on the original unoptimized table. This takes more than eight minutes. A query execution of such a high duration is far away from real time calculation and

% To be continued
